---
title: "tidyverse probs"
editor: visual
format: revealjs
---

## loading tidyverse




```{r}
#| echo: true
library(tidyverse)
library(tidybiology)
```

## loading dataframes

```{r}
#| echo: true
data(happy)
data(happy_join_one)
data(happy_full)
```

## taking a closer look at the data frames

```{r}
#| echo: true
glimpse(happy)
glimpse(happy_join_one)
glimpse(happy_full)
```

We know we can joing these data frames because they all share the column "ladder score."

## Combining "happy" and "happy_join_one" data frames by row

```{r}
#| echo: true

bind_rows(happy, happy_join_one)
```

## Checking to make sure that "happy" and "happy_join_one" data frames have equal numbers of rows

```{r}
#| echo: true

nrow(happy) == nrow(happy_join_one)
```

These two data frames do have equal numbers of rows

## Combining the datasets "happy" and "happy_join_one" by column

```{r}
#| echo: true

bind_cols(happy, happy_join_one)
```

These dataframes both have columns called "country_name."

## Joining the dataframes with a cleaner output

```{r}
#| echo: true

happy %>%
  inner_join(happy_join_one)
```
This function combines the dataframes so that they share a column labeled "country_names."

```{r}
#| echo: true

happy_full %>%
  anti_join(happy)
```

This function filters out all countries from "happy_full" that are also in "happy"